Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-07 ms.tex
conclusions 11
conclusions 12

Another approach to determine the SMBH accretion rate, which we adopt
in this paper, is to directly calculate the density, velocity and
temperature profiles of the CNM using a physically motivated
hydrodynamic model.  Mass is injected to the nuclear environment via
stellar winds, while energy input results from several sources
including stellar winds, supernovae, and AGN feedback
(\citealt{Quataert:2004a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}).
Unlike previous works, which focused primarily on modeling individual galaxies, here
we systematically analyze the CNM properties across a range of galaxy properties, including different SMBH masses, stellar density profiles, and
star formation histories (\citealt{WangMerritt:2004a}).

@@

The source term in equation (\ref{eq:drhodt}),
\begin{align}
  q=\frac{\eta \rhostar}{\th},
\label{eq:q}
\end{align}
represents mass input from stellar winds, which we parameterize in
terms of the fraction $\eta$ of the stellar density $\rhostar$ being
recycled into gas on the Hubble time $\th = 1.4 \times 10^{10}$ yr.
To good approximation $\eta\simeq 0.02 (\tau_{\star}/t_h)^{-1.3}$ at
time $\tau_{\star}$ following an impulsive star burst (e.g.,
\citealt{Ciotti+91})\footnote{\citet{Ciotti+91} give the mass return
  rate from evolved stars as a function of B-band luminosity instead
  of volumetrically, but our expressions are equivalent.}, although
$\eta$ is significantly higher for continuous star formation histories
(bottom panel of Fig.~\ref{fig:vwSources}).

@@

In Appendix \ref{app:windheat} we calculate the wind heating rate from
stellar winds, $v_{\rm w}^{\star}$, and the mass loss parameter,
$\eta$ (eq.~[\ref{eq:q}]), as a function of age, $\tau_{\star}$, of a
stellar population which is formed impulsively (Fig.~\ref{fig:vwImp}).  
At the earliest times ($\tau_{\star} \lesssim 10^{7}$ yr), the wind
heating rate exceeds 1000 km s$^{-1}$, while much later ($\tau_{\star}
\sim t_{\rm h}$) stellar wind heating dominated by main sequence winds
is much lower, $v^{\star}_{\rm w} \sim 50-100 $ km s$^{-1}$.  As shown below in $\S\ref{sec:combined}$ that for the case of quasi-continuous star
formation representative of the average star formation history (SFH) of loss
mass galaxies, the stellar wind heating rate produced by stellar winds
can also be significant, $v_{\rm w}^{\star} \sim 1000$ km s$^{-1}$.

@@

As in the case of Compton heating, kinetic heating could in principle
`self-regulate' the accretion flow insofar as a lower heating rate
results in a higher accretion rate (eq.~[\ref{eq:vjet}]), which in
turn may create stronger kinetic feedback.  However, given the uncertainty
in the efficiency of kinetic heating, we defer a more detailed discussion
to $\S\ref{sec:kinetic}$.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{vwSources.pdf}
\caption{\label{fig:vwSources} {\it Top Panel:} Sources contributing
  to the total heating rate of the CNM, $\vwO$ ({\it black}),
  including stellar wind heating ({\it orange}), Ia supernovae ({\it
    green}), millisecond pulsars ({\it blue}), and compton heating
  ({\it pink}).  Each heating source varies with black hole mass
  $\Mbh$ calculated for average star formation histories from
  \citet{MosterNaab+:2013a} (see Appendix~\ref{app:windheat} for
  details).  {\it Middle Panel:} The ratio of the total heating rate
  ($\dot{q}_{\rm heat} \propto v_{w}^{2}$) from the top panel to the
  radiating cooling rate ($|\dot{q}_{\rm rad}|$) at the stagnation
  radius. {\it Bottom panel} Parameter $\eta$ characterizing stellar
  mass loss rate (eq.~[\ref{eq:q}]) as a function of black hole mass
  $\Mbh$, calculated for average star SFHs from \citealt{MosterNaab+:2013a} and the
  ratio of Ia radius to stagnation radius as a function of $\Mbh$.}
  %The Ia radius calculated from equation~\eqref{eq:rIa} based on
  %convolving the average star formation histories with the SN Ia DTD.
  %The stagnation radius is calculated from \eqref{eq:stag_simple}
  %using the total heating rate for each $\Mbh$ shown in the top panel
  %assuming a cusp galaxy ($\Gamma=0.8$).}
\end{figure}

@@

The total external gas heating rate, 
\begin{equation}
\vwO = \sqrt{(v_{w}^{\star})^{2} + (v_{w}^{\rm MSP})^{2} + (v_{w}^{\rm Ia})^{2} + (v_{w}^{\bullet})^{2}},
\label{eq:vtot}
\end{equation}
includes contributions from stellar winds, supernovae, pulsars, and
radiative SMBH feedback.  The strength of each heating source depends
explicitly on the SMBH mass and the stellar population in the galactic
nuclear region.  As we now describe, the latter could best be
described by a single star burst formation in the past, or by a more
continuous star formation history that itself varies systematically
with the galaxy mass and hence $\Mbh$.

@@

Figure~\ref{fig:vwSourcesImp} shows the contributions of various
heating sources as a function of time $\tau_{\star}$ after a burst of
star formation for black holes of mass $\Mbh = 10^{6} \Msun$ (top
panel) and $\Mbh = 10^{8} \Msun$ (bottom panel). The heating and mass
input parameters due to stellar winds, $v_{w}^{\star}(\tau_{\star})$
and $\eta(\tau_{\star})$, are calculated as described in
Appendix~\ref{app:windheat} (Fig.~\ref{fig:vwImp}).  The heating from
SN Ia ($v_{w}^{\rm Ia}$) and Compton heating ($v_{\rm w}^{\rm C}$) are
calculated from equations~\eqref{eq:vIa} and~\eqref{eq:vC}
respectively, also making use of $\eta(\tau_{\star})$ from stellar
winds.  SN Ia heating depends also on a convolution of the DTD
(eq. [\ref{eq:DTD}]) and the star formation history, which reduces to
the DTD itself for a single star burst.

@@

\subsubsection{Continuous Star Formation History}
A single burst of star formation does not describe the typical star
formation history of most galaxies.  Qualitatively, smaller galaxies
will on average have more recent star formation, resulting in
energetic young stellar winds and supernovae which dominate the gas
heating budget.  Massive galaxies, on the other hand, will on average
possess older stellar populations, with their heating rates dominated
by SN Ia (on large radial scales) and SMBH feedback.  We estimate the
average value of $\vwO$ as a function of SMBH mass for each heating
source by calculating its value using the average cosmic star
formation histories of \citet{MosterNaab+:2013a}.  Appendix
\ref{app:windheat} describes how the average star formation history is
used to determine the stellar wind heating $v_{\rm w}^{\star}$ and
mass return parameter $\eta$ as a function of $M_{\bullet}$
(Fig.~\ref{fig:vwSources}; top panel).  Star formation histories are also
convolved with the Ia DTD distribution to determine the SN Ia heating.

Figure~\ref{fig:vwSources} shows $\vwO(M_{\bullet})$ from each heating
source (stars, MSPs, SNe Ia, and black hole feedback) calculated for
the average star formation history of galaxies containing a given
black hole mass.  The younger stellar populations characterizing galaxies
with $\Mbh\lsim 3\times 10^{8} \Msun$ are dominated by stellar winds, with $v_{w}^{\star} \gtrsim 700$ km s$^{-1}$.  Only for $\Mbh\gsim 3\times 10^{8} \Msun$ does the lack
of young stellar populations significantly reduce the role of stellar
wind heating.  But for these massive galaxies, that Ia radius is sufficiently small for Ia heating to contribute a comparable level of heating, $v_{w}^{\rm Ia} \gtrsim 600$ km s$^{-1}$ (for low mass galaxies Ia heating does not contribute to the total heating rate because $r_{\rm Ia} > \rs$; bottom panel).  Strikingly, we find that galactic nuclei experiencing the same
  average star formation history as their host galaxies produce thermally stable flows across almost all black hole masses.

@@

Realistic variations (``burstiness") in the star formation history
will also often produce lower stellar wind heating rates (Appendix
\ref{app:windheat}).  Figure \ref{fig:NickPlot2} shows the heating
rate as a function of average black hole mass for non-fiducial cases
in which the current ($z = 0$) star formation is suppressed by a
factor $\iota$ from its average $z = 0$ value, for a characteristic
timescale $\delta t_{\star}$.  For $\delta t_{\star} \lesssim 10^{7}$
yr, we see that $v_{\rm w}^{\star}$ is reduced by at most a factor of
$\approx 2$ from its average value, even for huge drops in the star
formation rate ($\iota \sim 10^{-3}$).  However, burstiness in the
star formation history over longer timescales can suppress heating
more significantly.  When $\delta t_{\star} \gtrsim 10^8~{\rm yr}$,
$v_{\rm w}^\star$ becomes very sensitive to $\iota$: at $\iota=0.1$,
$v_{\rm w}^\star \approx 500~{\rm km~s}^{-1}$ in most galaxies, but
for smaller $\iota$, $v_{\rm w}^\star \approx 200~{\rm km~s}^{-1}$ (an
exception to this is in galactic nuclei with $M_\bullet \gtrsim
5\times 10^8~M_\odot$, where heating is stabilized by SN Ia).  Our use
of average cosmic star formation histories is appropriate provided
local variations are either on short timescales ($\delta t_{\star}
\lesssim 10^7~{\rm yr}$), or limited in magnitude ($\iota \gtrsim
0.1$).  If both of these conditions are violated, then the effective
heating rates for all but the most massive galaxies (where Ia
explosions dominate) will fall by a factor $\approx 5$ from our
fiducial volumetric averages, taken from \citet{MosterNaab+:2013a}.

@@

Figure \ref{fig:vwSources} shows that the heating associated with the
average star formation history of a galaxy is approximately constant
with SMBH mass, except for the largest black holes with $M_{\bullet}
\gtrsim 5\times 10^{8}M_{\odot}$.  For a fixed wind heating parameter,
the Eddington ratio $\dot{M}/\dot{M}_{\rm edd}$ increases $\propto
M_{\bullet}^{0.5(0.8)}$ for core(cusp) galaxies, respectively
(Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]).

@@

\begin{figure*}
\includegraphics[width=\textwidth]{miller.pdf}
\caption{\label{fig:miller} Average nuclear X-ray luminosity, $\langle
  L_X \rangle = \epsilon_X \Mdot c^2$, as a function of SMBH mass.
  Black lines show our prediction in the case of heating due to the
  average star formation history for galaxies corresponding to each
  black hole mass (Fig.~\ref{fig:bh_growth}), calculated using a
  power-law fit for the relationship $\Gamma(M_{\bullet})$ derived
  from the \citet{LauerFaber+:2007a} sample.  Shown for comparison are
  the measurements ({\it red stars}) and upper limits ({\it red
    triangles}) for $L_X/L_{\rm edd}$ values, from the
  \citet{Miller+15} sample of early-type galaxies (a green line shows
  the best power-law fit $\langle L_X/L_{\rm edd}\rangle \propto
  M_{\bullet}^{\alpha}$, with $\alpha = -0.2$).  The left panel is
  calculated assuming a constant radiative efficiency
  $\epsilon_X=10^{-4} (\epsilon_{\rm bol}/0.1)$, while the right hand
  panel assumes $\epsilon_X= (\epsilon_{\rm
    bol}/0.1)(\Mdot/\MdotEdd)$, as predicted by ADAF models (e.g.,
  \citealt{Narayan&Yi95}).  Shown for comparison are the X-ray
  luminosities calculated for the maximum thermally-stable accretion
  rate ({\it orange line}; eq.~[\ref{eq:Mdotmax}]); the SN
  Ia-regulated accretion rate ({\it teal line};
  eq.~[\ref{eq:eddr_Ia}]), and the Compton heating-regulated accretion
  rate ({\it pink line}; eq.~[\ref{eq:MdotCa}]), again all calculated
  for the average star formation history corresponding to each SMBH
  mass.}
\end{figure*}

If the nuclei of elliptical galaxies are heated as expected for the average star formation history of galaxies with
similar mass, then to first order we predict that $\langle
L_{X}/L_{\rm edd}\rangle$ should be a weakly increasing function of BH mass.
This result is in tension with the observed down-sizing trend that the average Eddington ratio is seen to
decrease (albeit weakly) with $M_{\bullet}$, especially in the $\epsilon_X= (\epsilon_{\rm
  bol}/0.1)(\Mdot/\MdotEdd)$ model.  However, one must keep in mind the enormous uncertainty in calculating the luminosity of the accretion flow close to the black hole in a single waveband based on the feeding rate on larger scales.  If disk outflows carry away most of the infalling mass before it reaches X-ray producing radii (e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}), then the angular momentum of the infalling gas might also influence the X-ray luminosity, in a way that could depend systematically on the stellar population and hence $M_{\bullet}$.

@@

Figure \ref{fig:bh_growth} (top panel) also shows that $\dot{M}_{\rm
  C} > \dot{M}_{\rm TI}$ across all black hole masses.  This indicates that even if
Compton heating could act to stabilize the flow in a quasi-steady state (which, as already discussed, is unlikely), the magnitude of the heating is insufficient to produce a thermally stable accretion rate given conditions
corresponding to the average star formation history.

@@

We have calculated steady-state models for the hot gaseous circumnuclear media of quiescent galaxies, under the assumption that gas is supplied exclusively by stellar wind mass loss and heated by shocked stellar winds, supernovae and black hole feedback.  We compute solutions for a range of different black hole masses, heating rates, and observationally-motivated stellar density profiles.  Our numerical results (Table \ref{table:models}) are used to verify and calibrate analytic relationships (Appendix \ref{app:rs}, \ref{app:be}).  The latter are in turn used to explore systematically how the SMBH accretion rate varies with black hole mass and the galaxy's star formation history.  Our results for $\dot{M}(\Mbh)$ are compared with observed trends of the nuclear X-ray luminosities of quiescent SMBHs and low luminosity AGN.  They are also used to address key questions such as the thermal stability of the accretion flow.  Our conclusions are summarized as follows.

@@

\item Including the effects of heat transport by electron conduction results in only moderate changes to the key properties of the flow (e.g. stagnation radius, mass accretion rate, and thermal stability) for causal values of the conduction saturation parameter $\phi < 0.1$ (eq.~[\ref{eq:conduction}]).

\item Radiation has a more pronounced influence on the structure of the gas flow.  Cooling instabilities are likely to develop when the radiative cooling rate exceed the rate of external heating near the stagnation radius.  This condition is equivalent to the more commonly invoked criterion $t_{\rm cool} \gg t_{\rm ff}$ criterion (e.g.~\citealt{McCourt+12}; \citealt{Li&Bryan14a}).

  \item The location of the stagnation radius, and hence the SMBH accretion rate, is a sensitive function of the gas heating rate $\propto v_w^{2}$.  Quantitatively, $\rs\propto v_w^{-2}$, implying that $\Mdot\propto  v_w^{-3.8 (-2.4)}$ for core (cusp) galaxies.  However, $\Mdot$ can increase even more rapidly with decreasing $v_w$ for two reasons: (1) when $v_w$ is comparable or less than the stellar velocity dispersion ($\zeta < \zeta_c$; eq.~$[\ref{eq:zetacrit}]$), then gas remains bound to the stellar bulge and cannot produce a flow with a positive kinetic energy interior to the break radius of the stellar density profile (Appendix \ref{app:be}); (2) For low $v_w$, gas near the stagnation radius becomes thermally unstable, which based on previous numerical studies (e.g.,\citealt{Ciotti+10}) is found to result in time-dependent cycles of alternating high and low accretion states ($\S\ref{sec:cycle}$).

\item Stellar wind heating, supernovae, and AGN feedback depend explicitely on the SMBH mass as well as the star formation history in the galactic nuclear region.  A recent star burst of age $\lesssim 10$ Myr produces mass and energy injection dominated by young stellar winds ($v_w \sim 1000$ km s$^{-1}$), while for an exclusively old stellar population, mass input is dominated by AGB winds and heating is dominated by AGN feedback and supernovae.  Ongoing, continuous star formation presents a hybrid situation: energy input is usually dominated by young stars, while mass input is dominated by the old stellar population (Appendix \ref{app:windheat}).  Galactic nuclei with $\Mbh \lesssim 10^8\Msun$ that experience heating corresponding to their average star formation histories from cosmological simulations receive stellar wind heating of $v_w\sim 800$ km s$^{-1}$.

\item Type Ia SNe only provide a continuous heating source exterior to the radius $\rIa$ (eq.~[\ref{eq:RIa}]) where the time interval between subsequent SNe exceeds the local dynamical timescale.  The heating provided by the average star formation history results in $\rIa < \rs$, except for the most massive galaxies.  Nevertheless, even if other sources of heating are low enough that $\rIa > \rs$, the stagnation radius is unlikely to exceed $\rIa$ due to Ia heating.  SN Ia thus cap the SMBH accretion rate at the value $\dot{M}_{Ia}$ (eq.~[\ref{eq:MdotIa}]), except possibly in high mass elliptical galaxies with large break radii, for which $\zeta(v_w^{\rm Ia}) \gtrsim \zeta_c$.

\item Stellar wind heating due to the average star formation history is sufficient to result in thermally-stable, steady-state accretion.  The resulting average Eddington-normalized accretion rate is predicted to increase weakly with $M_{\bullet}$ (Fig.~\ref{fig:miller}), in tension with the (also weak) downsizing trend of measured nuclear X-ray luminosities of early-type galaxies (e.g., \citealt{Miller+15}).  However, given the large uncertainties in translating predictions for the gas inflow rate on parsec scales to the disk X-ray luminosity, these differences may not be significant.

>>>>

Another approach to determine the SMBH accretion rate, which we adopt
in this paper, is to directly calculate the density, velocity and
temperature profiles of the CNM using a physically motivated
hydrodynamic model.  Mass is injected to the nuclear environment via
stellar winds, while energy input results from several sources
including stellar winds, supernovae, and AGN feedback
(\citealt{Quataert:2004a,De-ColleGuillochon+:2012a,ShcherbakovWong+:2014a}).
Unlike previous works, which focused primarily on modeling individual galaxies, here
we systematically analyze the CNM properties across a range of galaxy properties, including different SMBH masses, stellar density profiles, and
star formation histories (SFHs; \citealt{WangMerritt:2004a}).

@@

The source term in equation (\ref{eq:drhodt}),
\begin{align}
  q=\frac{\eta \rhostar}{\th},
\label{eq:q}
\end{align}
represents mass input from stellar winds, which we parameterize in
terms of the fraction $\eta$ of the stellar density $\rhostar$ being
recycled into gas on the Hubble time $\th = 1.4 \times 10^{10}$ yr.
To good approximation $\eta\simeq 0.02 (\tau_{\star}/t_h)^{-1.3}$ at
time $\tau_{\star}$ following an impulsive star burst (e.g.,
\citealt{Ciotti+91})\footnote{\citet{Ciotti+91} give the mass return
  rate from evolved stars as a function of B-band luminosity instead
  of volumetrically, but our expressions are equivalent.}, although
$\eta$ is significantly higher for continuous SFHss
(bottom panel of Fig.~\ref{fig:vwSources}).

@@

In Appendix \ref{app:windheat} we calculate the wind heating rate from
stellar winds, $v_{\rm w}^{\star}$, and the mass loss parameter,
$\eta$ (eq.~[\ref{eq:q}]), as a function of age, $\tau_{\star}$, of a
stellar population which is formed impulsively (Fig.~\ref{fig:vwImp}).  
At the earliest times ($\tau_{\star} \lesssim 10^{7}$ yr), the wind
heating rate exceeds 1000 km s$^{-1}$, while much later ($\tau_{\star}
\sim t_{\rm h}$) stellar wind heating dominated by main sequence winds
is much lower, $v^{\star}_{\rm w} \sim 50-100 $ km s$^{-1}$.  As shown below in $\S\ref{sec:combined}$ that for the case of quasi-continuous star
formation representative of the average SFHs of low
mass galaxies, the stellar wind heating rate produced by stellar winds
can also be significant, $v_{\rm w}^{\star} \sim 1000$ km s$^{-1}$.

@@

As in the case of Compton heating, kinetic heating could in principle
`self-regulate' the accretion flow insofar as a lower heating rate
results in a higher accretion rate (eq.~[\ref{eq:vjet}]), which in
turn may create stronger kinetic feedback.  However, given the uncertainty
in the efficiency of kinetic heating, we hereafter exclude its effect and defer further discussion
to $\S\ref{sec:kinetic}$.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{vwSources.pdf}
\caption{\label{fig:vwSources} {\it Top Panel:} Sources contributing
  to the total heating rate of the CNM, $\vwO$ ({\it black}),
  including stellar wind heating ({\it orange}), Ia supernovae ({\it
    green}), millisecond pulsars ({\it blue}), and compton heating
  ({\it pink}).  Each heating source varies with black hole mass
  $\Mbh$ calculated for average SFHs from
  \citet{MosterNaab+:2013a} (see Appendix~\ref{app:windheat} for
  details).  {\it Middle Panel:} The ratio of the total heating rate
  ($\dot{q}_{\rm heat} \propto v_{w}^{2}$) from the top panel to the
  radiating cooling rate ($|\dot{q}_{\rm rad}|$) at the stagnation
  radius. {\it Bottom panel} Parameter $\eta$ characterizing stellar
  mass loss rate (eq.~[\ref{eq:q}]) as a function of black hole mass
  $\Mbh$, calculated for average star SFHs from \citealt{MosterNaab+:2013a} and the
  ratio of Ia radius to stagnation radius as a function of $\Mbh$.}
  %The Ia radius calculated from equation~\eqref{eq:rIa} based on
  %convolving the average star formation histories with the SN Ia DTD.
  %The stagnation radius is calculated from \eqref{eq:stag_simple}
  %using the total heating rate for each $\Mbh$ shown in the top panel
  %assuming a cusp galaxy ($\Gamma=0.8$).}
\end{figure}

@@

The total external gas heating rate, 
\begin{equation}
\vwO = \sqrt{(v_{w}^{\star})^{2} + (v_{w}^{\rm MSP})^{2} + (v_{w}^{\rm Ia})^{2} + (v_{w}^{\bullet})^{2}},
\label{eq:vtot}
\end{equation}
includes contributions from stellar winds, supernovae, pulsars, and
radiative SMBH feedback.  The strength of each heating source depends
explicitly on the SMBH mass and the stellar population in the galactic
nuclear region.  As we now describe, the latter could best be
described by a single star burst formation in the past, or by a more
continuous SFH that itself varies systematically
with the galaxy mass and hence $\Mbh$.

@@

Figure~\ref{fig:vwSourcesImp} shows the contributions of various
heating sources as a function of time $\tau_{\star}$ after a burst of
star formation for black holes of mass $\Mbh = 10^{6} \Msun$ (top
panel) and $\Mbh = 10^{8} \Msun$ (bottom panel). The heating and mass
input parameters due to stellar winds, $v_{w}^{\star}(\tau_{\star})$
and $\eta(\tau_{\star})$, are calculated as described in
Appendix~\ref{app:windheat} (Fig.~\ref{fig:vwImp}).  The heating from
SN Ia ($v_{w}^{\rm Ia}$) and Compton heating ($v_{\rm w}^{\rm C}$) are
calculated from equations~\eqref{eq:vIa} and~\eqref{eq:vC}
respectively, also making use of $\eta(\tau_{\star})$ from stellar
winds.  SN Ia heating depends also on a convolution of the DTD
(eq. [\ref{eq:DTD}]) and the SFH, which reduces to
the DTD itself for a single star burst.

@@

\subsubsection{Continuous SFH}
A single burst of star formation does not describe the typical star
formation history of most galaxies.  Qualitatively, smaller galaxies
will on average have more recent star formation, resulting in
energetic young stellar winds and supernovae which dominate the gas
heating budget.  Massive galaxies, on the other hand, will on average
possess older stellar populations, with their heating rates dominated
by SN Ia (on large radial scales) and SMBH feedback.  We estimate the
average value of $\vwO$ as a function of SMBH mass for each heating
source by calculating its value using the average cosmic star
formation histories of \citet{MosterNaab+:2013a}.  Appendix
\ref{app:windheat} describes how the average SFH is
used to determine the stellar wind heating $v_{\rm w}^{\star}$ and
mass return parameter $\eta$ as a function of $M_{\bullet}$
(Fig.~\ref{fig:vwSources}; top panel).  SFHs are also
convolved with the Ia DTD distribution to determine the SN Ia heating.

Figure~\ref{fig:vwSources} shows $\vwO(M_{\bullet})$ from each heating
source (stars, MSPs, SNe Ia, and black hole feedback) calculated for
the average SFH of galaxies containing a given
black hole mass.  The younger stellar populations characterizing galaxies
with $\Mbh\lsim 3\times 10^{8} \Msun$ are dominated by stellar winds, with $v_{w}^{\star} \gtrsim 700$ km s$^{-1}$.  Only for $\Mbh\gsim 3\times 10^{8} \Msun$ does the lack
of young stellar populations significantly reduce the role of stellar
wind heating.  But for these massive galaxies, that Ia radius is sufficiently small for Ia heating to contribute a comparable level of heating, $v_{w}^{\rm Ia} \gtrsim 600$ km s$^{-1}$ (for low mass galaxies Ia heating does not contribute to the total heating rate because $r_{\rm Ia} > \rs$; bottom panel).  Strikingly, we find that galactic nuclei experiencing the same
  average SFH as their host galaxies produce thermally stable flows across almost all black hole masses.

@@

Realistic variations (``burstiness") in the SFH
will also often produce lower stellar wind heating rates (Appendix
\ref{app:windheat}).  Figure \ref{fig:NickPlot2} shows the heating
rate as a function of average black hole mass for non-fiducial cases
in which the current ($z = 0$) star formation is suppressed by a
factor $\iota$ from its average $z = 0$ value, for a characteristic
timescale $\delta t_{\star}$.  For $\delta t_{\star} \lesssim 10^{7}$
yr, we see that $v_{\rm w}^{\star}$ is reduced by at most a factor of
$\approx 2$ from its average value, even for huge drops in the star
formation rate ($\iota \sim 10^{-3}$).  However, burstiness in the
SFH over longer timescales can suppress heating
more significantly.  When $\delta t_{\star} \gtrsim 10^8~{\rm yr}$,
$v_{\rm w}^\star$ becomes very sensitive to $\iota$: at $\iota=0.1$,
$v_{\rm w}^\star \approx 500~{\rm km~s}^{-1}$ in most galaxies, but
for smaller $\iota$, $v_{\rm w}^\star \approx 200~{\rm km~s}^{-1}$ (an
exception to this is in galactic nuclei with $M_\bullet \gtrsim
5\times 10^8~M_\odot$, where heating is stabilized by SN Ia).  Our use
of average cosmic SFHs is appropriate provided
local variations are either on short timescales ($\delta t_{\star}
\lesssim 10^7~{\rm yr}$), or limited in magnitude ($\iota \gtrsim
0.1$).  If both of these conditions are violated, then the effective
heating rates for all but the most massive galaxies (where Ia
explosions dominate) will fall by a factor $\approx 5$ from our
fiducial volumetric averages, taken from \citet{MosterNaab+:2013a}.

@@

Figure \ref{fig:vwSources} shows that the heating associated with the
average SFH of a galaxy is approximately constant
with SMBH mass, except for the largest black holes with $M_{\bullet}
\gtrsim 5\times 10^{8}M_{\odot}$.  For a fixed wind heating parameter,
the Eddington ratio $\dot{M}/\dot{M}_{\rm edd}$ increases $\propto
M_{\bullet}^{0.5(0.8)}$ for core(cusp) galaxies, respectively
(Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]).

@@

\begin{figure*}
\includegraphics[width=\textwidth]{miller.pdf}
\caption{\label{fig:miller} Average nuclear X-ray luminosity, $\langle
  L_X \rangle = \epsilon_X \Mdot c^2$, as a function of SMBH mass.
  Black lines show our prediction in the case of heating due to the
  average SFH for galaxies corresponding to each
  black hole mass (Fig.~\ref{fig:bh_growth}), calculated using a
  power-law fit for the relationship $\Gamma(M_{\bullet})$ derived
  from the \citet{LauerFaber+:2007a} sample.  Shown for comparison are
  the measurements ({\it red stars}) and upper limits ({\it red
    triangles}) for $L_X/L_{\rm edd}$ values, from the
  \citet{Miller+15} sample of early-type galaxies (a green line shows
  the best power-law fit $\langle L_X/L_{\rm edd}\rangle \propto
  M_{\bullet}^{\alpha}$, with $\alpha = -0.2$).  The left panel is
  calculated assuming a constant radiative efficiency
  $\epsilon_X=10^{-4} (\epsilon_{\rm bol}/0.1)$, while the right hand
  panel assumes $\epsilon_X= (\epsilon_{\rm
    bol}/0.1)(\Mdot/\MdotEdd)$, as predicted by ADAF models (e.g.,
  \citealt{Narayan&Yi95}).  Shown for comparison are the X-ray
  luminosities calculated for the maximum thermally-stable accretion
  rate ({\it orange line}; eq.~[\ref{eq:Mdotmax}]); the SN
  Ia-regulated accretion rate ({\it teal line};
  eq.~[\ref{eq:eddr_Ia}]), and the Compton heating-regulated accretion
  rate ({\it pink line}; eq.~[\ref{eq:MdotCa}]), again all calculated
  for the average SFH corresponding to each SMBH
  mass.}
\end{figure*}

If the nuclei of elliptical galaxies are heated as expected for the average SFH of galaxies with
similar mass, then to first order we predict that $\langle
L_{X}/L_{\rm edd}\rangle$ should be a weakly increasing function of BH mass.
This result is in tension with the observed down-sizing trend that the average Eddington ratio is seen to
decrease (albeit weakly) with $M_{\bullet}$, especially in the $\epsilon_X= (\epsilon_{\rm
  bol}/0.1)(\Mdot/\MdotEdd)$ model.  However, one must keep in mind the enormous uncertainty in calculating the luminosity of the accretion flow close to the black hole in a single waveband based on the feeding rate on larger scales.  If disk outflows carry away most of the infalling mass before it reaches X-ray producing radii (e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}), then the angular momentum of the infalling gas might also influence the X-ray luminosity, in a way that could depend systematically on the stellar population and hence $M_{\bullet}$.

@@

Figure \ref{fig:bh_growth} (top panel) also shows that $\dot{M}_{\rm
  C} > \dot{M}_{\rm TI}$ across all black hole masses.  This indicates that even if
Compton heating could act to stabilize the flow in a quasi-steady state (which, as already discussed, is unlikely), the magnitude of the heating is insufficient to produce a thermally stable accretion rate given conditions
corresponding to the average SFH.

@@

We have calculated steady-state models for the hot gaseous circumnuclear media of quiescent galaxies, under the assumption that gas is supplied exclusively by stellar wind mass loss and heated by shocked stellar winds, supernovae and black hole feedback.  We compute solutions for a range of different black hole masses, heating rates, and observationally-motivated stellar density profiles.  Our numerical results (Table \ref{table:models}) are used to verify and calibrate analytic relationships (Appendix \ref{app:rs}, \ref{app:be}).  The latter are in turn used to explore systematically how the SMBH accretion rate varies with black hole mass and the galaxy's SFH.  Our results for $\dot{M}(\Mbh)$ are compared with observed trends of the nuclear X-ray luminosities of quiescent SMBHs and low luminosity AGN.  They are also used to address key questions such as the thermal stability of the accretion flow.  Our conclusions are summarized as follows.

@@

\item Including the effects of heat transport by electron conduction results in at most order unity changes to the key properties of the flow (e.g. stagnation radius, mass accretion rate, and thermal stability) for causal values of the conduction saturation parameter $\phi < 0.1$ (eq.~[\ref{eq:conduction}]).  Angular momentum is also unlikely to invalidate our spherically symmetric models given the extrapolated rotation curves of galaxies to the nuclear scales of interest ($\S\ref{sec:rotation}$).

\item Radiative cooling has a more pronounced influence on the structure of the gas flow.  Cooling instabilities are likely to develop when the radiative cooling rate exceeds the gas heating rate near the stagnation radius.  This condition is approximately equivalent to the more criterion $t_{\rm cool} \gg t_{\rm ff}$ criterion advocated by \citet{McCourt+12} and \citet{Li&Bryan14a}.

  \item The location of the stagnation radius, and hence the SMBH accretion rate, is a sensitive function of the gas heating rate $\propto v_w^{2}$.  Quantitatively, $\rs\propto v_w^{-2}$, implying that $\Mdot\propto  v_w^{-3.8 (-2.4)}$ for core (cusp) galaxies.  However, $\Mdot$ can increase even more rapidly with decreasing $v_w$ for two reasons: (1) when $v_w$ becomes comparable to the stellar velocity dispersion ($\zeta < \zeta_c$; eq.~$[\ref{eq:zetacrit}]$), then gas remains bound to the stellar bulge and cannot produce a flow with a positive kinetic energy interior to the stellar break radius (Appendix \ref{app:be}); (2) For low $v_w$, gas near the stagnation radius becomes thermally unstable, which based on the results of previous numerical studies (e.g.,\citealt{Ciotti+10}) is instead likely to result in time-dependent cycles of alternating high and low accretion phases ($\S\ref{sec:cycle}$).

\item Stellar wind heating, supernovae, and AGN feedback depend explicitely on the SMBH mass as well as the SFH in the galactic nuclear region.  A young starburst of age $\lesssim 10$ Myr produces mass and energy injection dominated by young stellar winds ($v_w \sim 1000$ km s$^{-1}$), while for exclusively old stellar populations  mass input is dominated by AGB winds and heating is dominated by AGN feedback and supernovae.  Ongoing, continuous star formation presents a hybrid situation, with energy input usually dominated by young stars and mass input dominated by the old stars (Appendix \ref{app:windheat}).  Galactic nuclei with $\Mbh \lesssim 10^8\Msun$ heated according to their average SFHs (as derived from cosmological simulations) receive a stellar wind heating of $v_w\sim 800$ km s$^{-1}$, although this number can be somewhat lower if the star formation is intermittent (Fig.~\ref{fig:NickPlot2}).

\item Type Ia SNe only provide a continuous heating source exterior to the Ia radius $\rIa$ (eq.~[\ref{eq:rIa}]) where the time between subsequent SNe exceeds the dynamical timescale.  Average SFH heating results in $\rIa < \rs$, except for the most massive galaxies.  Nevertheless, even if other sources of heating are low enough that $\rIa > \rs$, the stagnation radius is unlikely to exceed $\rIa$ due to Ia heating.  SN Ia thus cap the SMBH accretion rate at the value $\dot{M}_{Ia}$ (eq.~[\ref{eq:MdotIa}]), except possibly in high mass elliptical galaxies with large break radii, for which $\zeta(v_w^{\rm Ia}) \gtrsim \zeta_c$.

\item Stellar wind heating due to the average SFH is sufficient to result in thermally-stable, steady-state accretion.  The resulting average Eddington-normalized accretion rate is predicted to increase weakly with $M_{\bullet}$ (Fig.~\ref{fig:miller}), in tension with the (also weak) downsizing trend of measured nuclear X-ray luminosities of early-type galaxies (e.g., \citealt{Miller+15}).  However, given the large uncertainties in translating predictions for the gas inflow rate on parsec scales to the disk X-ray luminosity, these differences may not be significant.

Thermally stable accretion models can reproduce scatter in nuclear X-ray luminosities at fixed $\Mbh$ of the observed 2-4 orders of magnitude from a combination of (1) intrinsic scatter in the SMBH scaling relations; (2) stochastic variation in SFH from galaxy to galaxy, or from time to time due to burstiness in the SFH (e.g.~\ref{fig:NickPlot2}); (3) Variations in the amount of inflowing mass which is lost to small scale outflows from the SMBH accretion disk, due to variations in the angular momentum of the accreting gas.
